\section{Introduction}

The preceding chapters of this thesis treated the Gromov--Wasserstein
(GW) distance as an object of study in its own right: a geometry-aware
divergence between metric--measure spaces \citep{memoli_2022_distance},
solved either exactly or through its entropic relaxation
\citep{peyre_2016_gromov, solomon_2016_entropic}, and used to align
representations \emph{post hoc} without involving a learned model.
This chapter changes the role of GW. Rather than measuring alignment
between two fixed embedding sets, we ask whether the same quantity is
useful as a \emph{differentiable regularizer} that shapes the
representations of a deep generative model \emph{during training}.

The setting we choose for this question is video-to-audio (V2A)
generation with flow matching \citep{lipman_2023_flow,
albergo_2023_building, liu_2023_rectified}, a now-standard
simulation-free training objective for continuous normalizing flows
that has rapidly become the dominant formulation for audio
generation \citep{guan_2024_lafma, wang_2024_frieren,
cheng_2025_mmaudio} and high-resolution multimodal synthesis
\citep{esser_2024_scaling}. V2A is a stress test for cross-modal
alignment: an audio waveform must be both \emph{semantically}
consistent with what is happening on screen and \emph{temporally}
synchronised with it. State-of-the-art V2A systems --- exemplified by MMAudio
\citep{cheng_2025_mmaudio}, the open-source flow-matching backbone
we build on --- handle this through a combination of pointwise
mechanisms: joint multimodal attention over concatenated
video/text/audio tokens, adaLN conditioning on a global vector $c_g$
broadcast to all tokens, and a frame-aligned conditioning $c_f$
driven by Synchformer features that modulates the audio stream
token-by-token. These mechanisms align the two modalities
\emph{point by point}: each audio token is pushed toward the video
token it should attend to, but the \emph{relations among} the audio
tokens are not explicitly required to mirror the relations among the
video tokens. In the language of optimal transport, the
training objective is shaped by Wasserstein-style (feature-level)
signals while the Gromov--Wasserstein-style (relational) signal is
left implicit, to be discovered by the network if and when it helps.

Recent work in adjacent settings suggests this is a missed
opportunity. \citet{chen_2020_graph} show that adding a joint
Wasserstein + Gromov--Wasserstein loss as a drop-in regularizer ---
sharing a single Sinkhorn-solved transport plan between the two
terms, weighted by a single mixing coefficient $\lambda$ ---
improves image--text retrieval, visual question answering, and
machine translation over attention-only baselines, with their
ablation favouring a relatively large weight on the Wasserstein term
while the Gromov--Wasserstein term contributes the structural lift. \citet{lu_2025_crossmodal}
report the same pattern when transferring linguistic knowledge into a
CTC speech recogniser: the GW edge-matching term yields gains
\emph{beyond} those of optimal-transport methods that match only
features. \citet{chandra_2025_realign} push the recipe to long
procedural videos and obtain double-digit F1 improvements over
Kantorovich-OT baselines using a fused partial GW loss.
\citet{gong_2022_gromov} go further still and show that pure GW
suffices to align modalities even when no paired data are available,
by matching their pairwise-distance structure alone.
\citet{sasaki_2025_unsupervised} formalise this perspective for
neuroscience and machine learning: GW recovers cross-system
correspondences --- between brains, species, or networks ---
that supervised representational similarity analysis flattens, because
it is invariant to isometric reparametrisations of each side.

What is conspicuously absent from this body of work is any use of GW
inside a \emph{generative} V2A objective. Existing cross-modal GW
results target discriminative tasks (retrieval, captioning, ASR,
clustering) or self-supervised pretraining; the closest generative
analogue we are aware of, \citet{wang_2024_inverse}'s
inverse-OT framework for music-guided trailer generation, uses only
the Wasserstein component. The question this chapter addresses is
therefore concrete: \emph{can a Gromov--Wasserstein regularizer that
aligns the relational geometry of video and audio embeddings improve
flow-matching V2A generation, beyond what pointwise conditioning
already achieves?}

We answer the question by adding a GW penalty to the standard
flow-matching loss of an open-source V2A backbone (MMAudio) trained on
VGGSound. The penalty is computed per mini-batch from intra-video and
intra-audio pairwise distance matrices and solved via the entropic
Sinkhorn relaxation in fp32, following \citet{peyre_2016_gromov} and
\citet{solomon_2016_entropic}. Two design choices distinguish our
implementation from a naive port of off-the-shelf GW. First, we
investigate four representation variants over which the pairwise
distances are computed --- using global pooled features, a learned
projection, the conditioning signal $c_g$, and a fused
Wasserstein--GW objective in the spirit of \citet{vayer_2019_optimal,
vayer_2019_fused}. Second, we schedule the regularization weight
$\lambda_{\mathrm{GW}}$ during training rather than fixing it,
allowing the relational signal to engage gradually as the
flow-matching loss settles. The four representation variants map
naturally onto MMAudio's existing conditioning signals: the
\texttt{c\_g} variant computes pairwise distances over the global
adaLN vector, \texttt{global} and \texttt{projected} act on the
visual / audio token streams, and \texttt{fused} combines a
Wasserstein term with the GW term in the spirit of GOT
\citep{chen_2020_graph}, sharing a single Sinkhorn transport plan
between the two.

The use of entropic GW as a per-batch loss is justified by the
recently established statistical theory:
\citet{zhang_2024_gromov} show that entropic GW achieves the
parametric $n^{-1/2}$ sample-complexity rate independent of the
ambient dimension, and \citet{rioux_2023_entropic} prove convergence
of the entropic-GW solver together with a central-limit theorem for
the empirical estimator. The remaining design tension --- the cubic
per-iteration cost of standard entropic-GW versus the linear-time
low-rank variant of \citet{scetbon_2022_lineartime} or the sliced
variant of \citet{vayer_2019_sliced}, and the equal-mass assumption
versus the unbalanced formulation of \citet{sejourne_2021_unbalanced}
--- is taken up in the methodology and discussion sections.

Concretely, this chapter contributes:
\begin{itemize}
  \item the first integration, to our knowledge, of a
        Gromov--Wasserstein regularizer into a flow-matching V2A
        objective, building on the entropic GW machinery introduced
        in earlier chapters;
  \item a four-way ablation across representation variants
        (\texttt{global}, \texttt{projected}, \texttt{c\_g},
        \texttt{fused}) that isolates which embedding geometry
        carries the cross-modal relational signal;
  \item a schedule for $\lambda_{\mathrm{GW}}$ that resolves an
        engineering question left open by prior fused-OT work, where
        the trade-off parameter is typically held fixed;
  \item an empirical evaluation on VGGSound that asks whether the
        relational signal contributes any improvement on top of the
        already strong pointwise conditioning of a modern V2A
        backbone.
\end{itemize}

The remainder of this chapter is organised as follows. Section
\ref{sec:related_work} situates the contribution among prior
cross-modal OT/GW work and modern V2A systems. Section
\ref{sec:methodology} introduces the flow-matching backbone, defines
the four GW variants and the $\lambda_{\mathrm{GW}}$ schedule, and
states the design axes (D1--D5) and hypotheses (H1--H5) that the
experiments are built to test. Sections \ref{sec:results} and
\ref{sec:discussion} report and interpret the resulting evidence.
