\section{Methodology}
\label{sec:methodology}

This section formalises the integration of a Gromov--Wasserstein
regularizer into a conditional flow-matching objective. We first
recall the flow-matching loss as instantiated in MMAudio
\citep{cheng_2025_mmaudio}, then define the four families of
intra-modal pairwise-distance matrices that feed the regularizer,
the entropic and fused-entropic Gromov--Wasserstein objectives, the
Sinkhorn solver used to evaluate them, and the schedule that
controls their contribution to training. Throughout, we treat
MMAudio purely as a conditioning architecture: a means of
materialising a video-conditioned generative process whose
representations the GW loss can shape. The thesis claim is about
the regularizer; the backbone is a substrate.

\subsection{Backbone objective: conditional flow matching}
\label{sec:methodology:fm}

\paragraph{From CNFs to flow matching.}
A continuous normalising flow (CNF) parameterises a generative model
as the time-1 map of an ordinary differential equation
$\dot x_t = v_\theta(t, x_t)$, transporting samples
$x_0 \sim p_0$ from a tractable source (typically
$p_0 = \mathcal{N}(0, I)$) to samples $x_1 \sim p_1$ from the data
distribution. Classical maximum-likelihood training of CNFs
requires backpropagating through an ODE solver, which is
prohibitively expensive at the scales of interest here. Flow
matching \citep{lipman_2023_flow}, developed concurrently as
*stochastic interpolants*
\citep{albergo_2023_building, albergo_2023_stochastic}, sidesteps
this cost: rather than maximising the likelihood induced by the
ODE, one fixes a *conditional probability path*
$p_t(x \mid x_1)$ between $p_0$ and $p_1$, derives its target
velocity field $u_t(x \mid x_1)$ in closed form, and trains
$v_\theta$ by direct $L_2$ regression onto the target. Training is
then *simulation-free*: each gradient step only requires sampling
$(t, x_0, x_1)$ and a single forward pass through $v_\theta$, with
no SDE or ODE integration in the loop. Inference still solves an
ODE, but only at sampling time.

\paragraph{The OT / linear path.}
Among admissible probability paths, the one induced by
optimal-transport displacement interpolation between $p_0$ and a
single data point $x_1$ has a particularly simple form: the
deterministic linear interpolation
\begin{equation}
  x_t \;=\; t\,x_1 + (1-t)\,x_0,
  \qquad x_0 \sim \mathcal{N}(0, I), \;\; t \sim p_t,
  \label{eq:fm_path}
\end{equation}
with the constant target velocity
$u(x_t \mid x_0, x_1) = x_1 - x_0$. This linear path is the limit
of the Gaussian-path family at zero variance and underlies the
\emph{rectified flow} formulation of \citet{liu_2023_rectified};
its straightness makes the learned ODE accurately integrable in
few Euler steps. \citet{tong_2024_improving} extend the recipe to
non-Gaussian sources by coupling $(x_0, x_1)$ with a minibatch OT
plan, recovering the same loss with an empirically straighter
flow. We adopt the simplest variant — Gaussian source, independent
coupling, linear path — matching the rectified-flow choice of the
backbone we build on.

\paragraph{Conditional flow matching for V2A.}
For video-conditioned audio generation, let
$x_1 \in \R^{N \times d_a}$ denote a target audio latent of $N$
tokens of dimension $d_a$ (in our default configuration $N=250$ at
$31.25$ fps and $d_a=20$, the VAE latent space inherited from
MMAudio's S-16kHz model). Let
$\mathbf{C} = (F_v, F_{\mathrm{syn}}, F_t)$ collect the conditioning
features extracted by frozen encoders: CLIP visual features $F_v$,
Synchformer features $F_{\mathrm{syn}}$, and CLIP text features
$F_t$. Conditional flow matching learns a time-dependent velocity
field
$v_\theta(t, \mathbf{C}, x_t) : [0,1] \times \R^{|\mathbf{C}|}
\times \R^{N\times d_a} \to \R^{N \times d_a}$
that pushes the Gaussian prior toward the data distribution along
the linear path \eqref{eq:fm_path}, conditioned on $\mathbf{C}$.
The flow-matching loss is then
\begin{equation}
  \LFM(\theta)
  \;=\;
  \E_{t, x_0, (x_1, \mathbf{C})}
  \bigl\lVert v_\theta(t, \mathbf{C}, x_t) - (x_1 - x_0) \bigr\rVert^{2}.
  \label{eq:fm_loss}
\end{equation}
Crucially, $\LFM$ is a \emph{token-wise} regression: the squared
error is averaged independently over the $N$ audio tokens, and the
target $x_1 - x_0$ for each token depends only on that token's
noise sample. Cross-modal information enters exclusively through
$\mathbf{C}$, and only via the learned velocity field. Whatever
relational structure aligns the audio tokens to the video tokens
must therefore be absorbed by $v_\theta$'s parameters --- there is
no term in \eqref{eq:fm_loss} that explicitly compares the geometry
of the two streams.

\paragraph{The MMAudio realisation.}
MMAudio \citep{cheng_2025_mmaudio} instantiates $v_\theta$ as a
stack of $N_1$ multimodal MM-DiT blocks
\citep{esser_2024_scaling} (visual / text / audio streams
communicating via joint attention) followed by $N_2$ audio-only
transformer blocks. Conditioning is injected at two levels:
adaptive layer-norm modulation by a global vector $c_g$ pooled
from the projected visual and text features, and token-level
modulation by a frame-aligned vector $c_f$ derived from
$F_{\mathrm{syn}}$ at $24$ fps. It reaches state of the art among
public V2A systems \citep{cheng_2025_mmaudio}, surpassing earlier
flow-matching V2A work such as Frieren
\citep{wang_2024_frieren} (channel-level fusion + reflow). None of
these mechanisms align the pairwise relations among audio tokens
to those among video tokens; each token of one stream is nudged
toward the correct token of the other stream by attention and
adaLN, but the geometry of the streams is left to emerge
incidentally from the token-wise regression in \eqref{eq:fm_loss}.
The regularizer below targets exactly this gap.

\subsection{Representation variants and the relational signal}
\label{sec:methodology:reps}

For a mini-batch of size $B$, write
$F_v^{(b)} \in \R^{n_v \times d_v}$ for the raw CLIP feature
sequence of sample $b$ (with $n_v$ visual tokens of dimension
$d_v=1024$) and $x_1^{(b)} \in \R^{N \times d_a}$ for the
corresponding audio latent. The regularizer takes as input two
families of squared-Euclidean intra-modal distance matrices, one
per side. We expose four ways of constructing them, each
corresponding to a different geometric hypothesis about \emph{which}
representations should align relationally.

The first three variants compress each token sequence into a single
vector and compute pairwise distances \emph{across} the batch,
yielding $B \times B$ matrices $\DV, \DA \in \R^{B \times B}$. The
fourth keeps token sequences intact and computes pairwise distances
\emph{within} each sample, yielding per-sample $n \times n$ and
$m \times m$ matrices. Concretely:

\begin{description}
  \item[\texttt{global}] uses pre-projection averages
    $\bar v^{(b)} = \tfrac{1}{n_v}\sum_i F_v^{(b),i}$ and
    $\bar x^{(b)} = \tfrac{1}{N}\sum_i x_1^{(b),i}$. This probes
    whether the raw encoder geometry already carries cross-modal
    relational structure, before any of MMAudio's learnable
    parameters intervene.
  \item[\texttt{projected}] passes the raw features through
    MMAudio's input projection layers $\pi_v, \pi_a$ before
    averaging:
    $\tilde v^{(b)} = \tfrac{1}{n_v}\sum_i \pi_v(F_v^{(b)})_i$,
    analogously for audio. The relational signal now back-propagates
    into $\pi_v$ and $\pi_a$, the first learnable layers the GW
    objective can reshape.
  \item[\texttt{c\_g}] uses the same audio side as
    \texttt{projected} but, on the video side, applies the global
    conditioning projection
    $c_g^{(b)} = \pi_g\bigl(\tfrac{1}{n_v}\sum_i \pi_v(F_v^{(b)})_i\bigr)$,
    i.e.\ the very vector that drives MMAudio's adaLN modulation.
    Aligning the geometry of $c_g$ to the audio geometry directly
    targets the conditioning signal that already flows through every
    transformer block.
  \item[\texttt{fused}] uses the same averaged projected
    representations as \texttt{projected}, but couples the GW
    objective to a Wasserstein term on the cross-modal cosine
    distance, yielding a fused-Gromov--Wasserstein loss in the
    sense of \citet{vayer_2019_optimal,vayer_2019_fused}. The
    transport plan is shared between the two terms, in the manner
    of \citet{chen_2020_graph}.
  \item[\texttt{got\_token}] keeps the per-sample token graphs.
    For each sample $b$ we form
    $D_{\mathrm{V}}^{(b)} \in \R^{n_v \times n_v}$ from the
    projected visual tokens and
    $D_{\mathrm{A}}^{(b)} \in \R^{N \times N}$ from the projected
    audio tokens, with a per-sample cross-modal cost
    $C_{\mathrm{xy}}^{(b)} \in \R^{n_v \times N}$, and solve a
    fused-GW problem \emph{per sample}. This is the formulation
    closest in spirit to GOT \citep{chen_2020_graph}, lifted from
    image--text token sets to per-sample video / audio token sets.
\end{description}

For all variants, raw squared distances $D$ are normalised by
their mean,
\begin{equation}
  \widehat D \;=\; D \,/\, \overline{D},
  \qquad
  \overline{D} \;=\; \mathrm{mean}(D),
  \label{eq:dist_normalisation}
\end{equation}
with the divisor detached from the autograd graph. This makes the
regularizer scale-invariant in the sense that multiplying any
representation by a constant leaves $\widehat D$ unchanged, which in
turn makes the trade-off coefficient $\lambda_{\mathrm{GW}}$
introduced in \S\ref{sec:methodology:training} comparable across
variants and across training steps even as feature magnitudes
drift.

\subsection{Entropic Gromov--Wasserstein}
\label{sec:methodology:egw}

Let $\widehat\DV, \widehat\DA \in \R^{B \times B}$ denote the
normalised intra-video and intra-audio distance matrices for one of
the cross-batch variants above, and equip both sides with the
uniform marginal
$p = q = \tfrac{1}{B}\mathbf{1}_B$. The quadratic
Gromov--Wasserstein discrepancy with squared cost is
\begin{equation}
  \mathrm{GW}^{2}(\widehat\DV, \widehat\DA, p, q)
  \;=\;
  \min_{T \in \Pi(p,q)}
  \sum_{i,j,k,l}
  \bigl(\widehat\DV_{ik} - \widehat\DA_{jl}\bigr)^{2}
  \, T_{ij} T_{kl},
  \label{eq:gw}
\end{equation}
where $\Pi(p,q) = \{T \in \R_{\geq 0}^{B \times B} :
T \mathbf{1}_B = p,\, T^{\!\top}\!\mathbf{1}_B = q\}$. Equation
\eqref{eq:gw} is a quadratic assignment problem and NP-hard in
general; following \citet{peyre_2016_gromov} and
\citet{solomon_2016_entropic} we work with its entropic relaxation
\begin{equation}
  \mathrm{EGW}_\varepsilon
  \;=\;
  \min_{T \in \Pi(p,q)}
  \;\bigl\langle L(\widehat\DV, \widehat\DA) \otimes T,\, T\bigr\rangle
  \;-\; \varepsilon\, H(T),
  \label{eq:egw}
\end{equation}
with $L(\widehat\DV, \widehat\DA)_{ijkl} = (\widehat\DV_{ik} -
\widehat\DA_{jl})^{2}$ and $H(T) = -\sum_{ij} T_{ij}(\log T_{ij} -
1)$ the discrete entropy. The entropic relaxation is convex in $T$
\emph{conditional on the linearisation around the current iterate},
and admits the Sinkhorn-iteration solver below; its statistical
behaviour at finite $B$ is governed by the parametric
$B^{-1/2}$ rate of \citet{zhang_2024_gromov} and the convergence
guarantees of \citet{rioux_2023_entropic}, neither of which holds
for the un-regularised problem.

\paragraph{Mirror-descent Sinkhorn solver.}
The entropic-GW solver follows the mirror-descent / Frank--Wolfe
recipe of \citet{peyre_2016_gromov} (introduced for GW barycenter
computation) and \citet{solomon_2016_entropic} (provably
convergent on entropic-GW correspondence problems). The idea is
to alternate two cheap steps: a \emph{linearisation} of the
quartic GW interaction around the current plan, and a single
\emph{entropic optimal-transport projection} onto the marginal
polytope solved by Sinkhorn iteration.

For squared cost the GW objective in \eqref{eq:gw} admits the
linearised gradient
\begin{equation}
  G(T) \;=\; -2\, \widehat\DV \, T \, \widehat\DA
  \;\;+\;\; \text{constants independent of } T,
  \label{eq:gw_gradient}
\end{equation}
since
$\partial / \partial T \bigl\langle L(\widehat\DV, \widehat\DA)
\otimes T, T\bigr\rangle = 2 (\widehat\DV^{2} p \mathbf{1}^{\!\top}
+ \mathbf{1} q^{\!\top} \widehat\DA^{2}) - 2 \widehat\DV T
\widehat\DA$, and the marginal-dependent terms are constant under
fixed $p, q$. At each outer iteration we therefore solve an
entropic optimal transport problem with cost matrix $G(T)$, that is,
\begin{equation}
  T^{+}
  \;=\;
  \arg\min_{T' \in \Pi(p,q)}
  \langle G(T), T'\rangle - \varepsilon H(T'),
  \label{eq:inner_eot}
\end{equation}
whose solution has the Gibbs form $T^{+}_{ij} = u_i K_{ij} v_j$
with $K = \exp(-G(T)/\varepsilon)$, and $u, v$ are recovered by the
Sinkhorn fixed-point recursion
\begin{equation}
  u \leftarrow p \,/\, (K v),
  \qquad
  v \leftarrow q \,/\, (K^{\!\top} u),
  \label{eq:sinkhorn}
\end{equation}
iterated until the marginals match to a tolerance fixed by the
inner iteration count. The recursion is the standard Bregman /
matrix-scaling iteration that defines the entropic-OT solution
and inherits its linear convergence in the marginal-error
metric; we treat its hyperparameters
(\texttt{inner\_sinkhorn\_iter}, $\varepsilon$) as fixed budgets
rather than tuning to a stopping criterion. Convergence of the
overall outer-inner scheme on entropic GW with quadratic cost is
guaranteed by \citet{rioux_2023_entropic}, who prove
that the entropic-GW objective is convex in the linearised
regime and admits provable rates --- the theoretical underpinning
that justifies running only a small number of outer steps in
practice.

We run \texttt{num\_sinkhorn\_iter}$=5$ outer mirror-descent
steps, each containing \texttt{inner\_sinkhorn\_iter}$=20$
Sinkhorn passes, with $\varepsilon = 0.1$, and we stabilise the
kernel with the standard log-sum-exp shift
$K = \exp(-G/\varepsilon - \max(-G/\varepsilon))$ that
\citet{feydy_2019_interpolating} popularised in their
GPU-friendly Sinkhorn implementation. The full computation is
wrapped in an \texttt{autocast(enabled=False)} block so that the
entire solver runs in \texttt{float32}, which we found necessary
to avoid overflow in $K$ when $\varepsilon$ is small or distances
are poorly scaled --- a known sensitivity of entropic-OT solvers
addressed in the same line of work
\citep{feydy_2019_interpolating, rioux_2023_entropic}.

\paragraph{Loss value.}
At the converged plan $T^{\star}$ we report the full quadratic
objective rather than the linearised gradient. With uniform
marginals the squared-cost expansion reads
\begin{equation}
  \LGW
  \;=\;
  \langle \widehat\DV^{\odot 2}, p p^{\!\top} \rangle
  \;+\;
  \langle \widehat\DA^{\odot 2}, q q^{\!\top} \rangle
  \;-\;
  2\,\langle \widehat\DV\, T^{\star}\, \widehat\DA,\, T^{\star} \rangle,
  \label{eq:gw_value}
\end{equation}
where $D^{\odot 2}$ denotes element-wise square. The first two
terms are constants in $T^{\star}$ but vary across mini-batches and
across training steps; reporting them keeps loss magnitudes
comparable across the four variants.

\subsection{Fused Gromov--Wasserstein}
\label{sec:methodology:fgw}

For the \texttt{fused} and \texttt{got\_token} variants we add a
linear (Wasserstein) cross-modal term and share a single transport
plan between the two contributions, following GOT
\citep{chen_2020_graph} and the FGW formulation of
\citet{vayer_2019_optimal, vayer_2019_fused}.

\paragraph{Cross-modal cosine cost.}
The pure-GW objective \eqref{eq:gw_value} compares
\emph{intra-modal pairwise distances} only; it is invariant to any
isometric reparametrisation of either modality, including
arbitrary global rotations or reflections. The fused term breaks
this invariance by re-introducing a direct video-to-audio
\emph{pointwise} comparison through the cross-modal cost matrix
$C_{\mathrm{xy}}$. Concretely, write
$\tilde v, \tilde a \in \R^{\bullet \times d}$ for the projected
representations on the video and audio sides --- where $\bullet$
is the batch index $b \in \{1,\dots,B\}$ for the cross-batch
\texttt{fused} variant, and the within-sample token index for
\texttt{got\_token}. We first $L_2$-normalise each row so that
every representation lies on the unit sphere of $\R^d$,
\begin{equation}
  \hat v_i \;=\; \tilde v_i \,/\, \lVert \tilde v_i \rVert_2,
  \qquad
  \hat a_j \;=\; \tilde a_j \,/\, \lVert \tilde a_j \rVert_2,
  \label{eq:l2norm}
\end{equation}
and then compute the matrix of pairwise cosine \emph{distances}
\begin{equation}
  \bigl[C_{\mathrm{xy}}\bigr]_{ij}
  \;=\;
  1 \;-\; \hat v_i^{\!\top} \hat a_j
  \;\in\; [0, 2].
  \label{eq:ccross}
\end{equation}
For the cross-batch \texttt{fused} variant this yields
$C_{\mathrm{xy}} \in \R^{B \times B}$ whose diagonal entries are
the cosine distances between the projected video and audio
representations of the \emph{same} sample $b$ --- the natural
"pointwise" cost that an OT plan with diagonal mass would pay zero
under perfect cross-modal feature alignment. For
\texttt{got\_token} the same construction is performed
\emph{within each sample} on the projected token sequences,
yielding a per-sample $C_{\mathrm{xy}}^{(b)} \in \R^{n \times m}$
that compares each video token to each audio token of the same
clip. Three choices deserve note. (i) The cosine form is preferred
to a Euclidean distance because $\tilde v$ and $\tilde a$ live in
different feature spaces with potentially very different scales:
normalising to the unit sphere removes the scale ambiguity and
keeps $C_{\mathrm{xy}} \in [0, 2]$ for any $d$ and any feature
magnitude, so $C_{\mathrm{xy}}$ and the mean-normalised intra-modal
costs $\widehat\DV, \widehat\DA$ live on comparable scales without
further rescaling. (ii) The construction reuses the projection
layers $\pi_v, \pi_a$ that already produce the intra-modal
representations, so the FGW objective adds no new learnable
parameters: the same projector outputs feed the GW and Wasserstein
terms, ensuring they pull on the same representation rather than on
two disjoint heads. (iii) The video side of $C_{\mathrm{xy}}$
inherits the asymmetric \texttt{detach\_video} treatment of
\S\ref{sec:methodology:training}, so the Wasserstein term too acts
as a one-way pressure on the audio representation.

\paragraph{Linearised gradient and loss.}
With $C_{\mathrm{xy}}$ in hand and mixing parameter
$\alpha \in [0,1]$, the linearised gradient at iterate $T$ becomes
\begin{equation}
  G_{\mathrm{FGW}}(T)
  \;=\;
  \alpha \cdot \bigl(-2\, \widehat\DV \, T \, \widehat\DA\bigr)
  \;+\;
  (1 - \alpha) \cdot C_{\mathrm{xy}},
  \label{eq:fgw_gradient}
\end{equation}
plugged into the same inner entropic-OT solve
\eqref{eq:inner_eot}--\eqref{eq:sinkhorn}. The reported loss is
the same convex combination evaluated at $T^{\star}$,
\begin{equation}
  \LFGW
  \;=\;
  \alpha \cdot \LGW(T^{\star})
  \;+\;
  (1 - \alpha) \cdot \langle C_{\mathrm{xy}},\, T^{\star} \rangle.
  \label{eq:fgw_value}
\end{equation}
\textit{Sharing} the plan across the two terms is non-trivial: it
forces the Wasserstein and Gromov--Wasserstein contributions to
agree on a single coupling, rather than letting each pull the
representations toward its own optimum. \citet{chen_2020_graph}
report empirically that the shared-plan variant outperforms two
independent plans on every cross-domain task they examine; we
adopt the same design.

\subsection{Anti-collapse penalty}
\label{sec:methodology:anticollapse}

A practical risk of pure GW alignment between cross-batch
representations is that the projector(s) collapse the batch onto
a single direction, sending $\widehat\DV$ and $\widehat\DA$ both
toward zero and trivially minimising \eqref{eq:gw_value}. This is
a structural pathology of any similarity-driven objective with a
learnable feature extractor and no negative-pair pressure --- the
same failure mode that motivated the design of non-contrastive
self-supervised methods such as Barlow Twins
\citep{zbontar_2021_barlow} (cross-correlation matrix toward
identity), VICReg \citep{bardes_2022_vicreg} (per-dimension
variance and off-diagonal covariance penalties), and W-MSE
\citep{ermolov_2021_whitening} (hard whitening of batch
embeddings). At a higher level it also resembles the
entropic-bias collapse of plain entropic-OT losses that motivated
the Sinkhorn-divergence construction of
\citet{feydy_2019_interpolating}. We adopt the **spectral
log-determinant** form of the regularizer, introduced by
\citet{ozsoy_2022_corinfomax} as the second-order
mutual-information surrogate underlying the CorInfoMax objective.
Concretely, we add a spectral penalty on the centred Gram matrix
\begin{equation}
  K(v) \;=\; \frac{1}{B-1}\,\bigl(v - \bar v\bigr)\bigl(v - \bar v\bigr)^{\!\top}
  \;\in\; \R^{B \times B},
  \label{eq:gram}
\end{equation}
where $v \in \R^{B \times d}$ stacks the batch representations,
and define
\begin{equation}
  \mathcal{L}_{\mathrm{AC}}(v)
  \;=\;
  -\log\det\bigl(K(v) + \eta I_B\bigr).
  \label{eq:anticollapse}
\end{equation}
The non-zero spectrum of the Gram $K(v)$ coincides with that of the
$d \times d$ feature covariance up to the $(d - B)\log\eta$
constant when $d > B$, so the Gram form is computationally
preferable when $d \gg B$ but penalises the same rank deficiency
as the covariance form of \citet{ozsoy_2022_corinfomax}.
A rank-1 collapse drives $B-1$ eigenvalues of $K$ to zero, sending
\eqref{eq:anticollapse} toward $(B-1)\,(-\log\eta)$ — a large
positive number that repels the optimiser. Compared to VICReg's
per-dimension variance hinge or Barlow Twins' off-diagonal squared
penalty, the log-det form is a \emph{soft barrier on the joint
spectrum} rather than a per-axis or per-pair constraint: it forbids
rank deficiency without dictating the anisotropy of the surviving
spectrum, which leaves the GW objective free to choose the
direction of variation. This is the same trade-off that
distinguishes soft whitening (CorInfoMax-style) from hard
whitening (W-MSE-style \citep{ermolov_2021_whitening}); we prefer
the soft form because the GW geometry is precisely what we want to
shape, not constrain. Critically, the penalty is applied to the
\emph{non-detached} representations, so its gradient always reaches
the projector even when the GW term is computed with the video
side detached (\S\ref{sec:methodology:training}).

\subsection{Training integration and the \texorpdfstring{$\lambda_{\mathrm{GW}}$}{lambda\_GW} schedule}
\label{sec:methodology:training}

The total loss optimised at training step $s$ is
\begin{equation}
  \mathcal{L}_{\mathrm{total}}(\theta; s)
  \;=\;
  \LFM(\theta)
  \;+\;
  \lambda_{\mathrm{GW}}(s) \cdot \mathcal{L}_{\mathrm{rel}}(\theta)
  \;+\;
  \lambda_{\mathrm{AC}} \cdot \mathcal{L}_{\mathrm{AC}}(\theta),
  \label{eq:total_loss}
\end{equation}
where $\mathcal{L}_{\mathrm{rel}} \in \{\LGW, \LFGW\}$ is selected
by the variant. Three implementation choices deserve emphasis.

\paragraph{Subset selection.}
The regularizer is computed only over the subset of the mini-batch
for which a real video is present (\texttt{video\_exist} mask),
since for audio-only samples MMAudio's own training procedure
replaces the visual stream with learnable empty tokens; aligning
audio geometry to the geometry of constants is meaningless. If
fewer than \texttt{min\_batch}$=4$ real videos remain in the
batch, the GW term is set to zero for that step.

\paragraph{Asymmetric gradient flow.}
By default we set \texttt{detach\_video}$=\mathtt{True}$, which
stops the gradient on the video side of $\widehat\DV$ before the GW
solver. The video encoder (CLIP / Synchformer) is frozen in any
case, but the projection layers $\pi_v$ and $\pi_g$ are not, and
without the detach the GW loss can satisfy itself by reshaping the
visual conditioning rather than the audio representation. Detaching
ensures that the GW term acts as a one-way pressure: the audio side
is asked to mirror the video geometry, not the other way round.
The anti-collapse penalty bypasses this detach so it can still
discipline the projector outputs.

\paragraph{Schedule.}
The coefficient $\lambda_{\mathrm{GW}}(s)$ is governed by a small
schedule family:
\begin{equation}
  \lambda_{\mathrm{GW}}(s)
  \;=\;
  \begin{cases}
    \lambda_{0} \cdot s / s_{\mathrm{w}}
      & \text{if } s < s_{\mathrm{w}} \;\text{and schedule}=\textsc{linear-rampup}, \\[2pt]
    0
      & \text{if } s < s_{\mathrm{w}} \;\text{and schedule}=\textsc{constant}, \\[2pt]
    \lambda_{0}
      & \text{if } s \geq s_{\mathrm{w}} \;\text{and schedule}\in\{\textsc{constant}, \textsc{linear-rampup}\}, \\[2pt]
    \tfrac{\lambda_{0}}{2}\bigl(1 + \cos\bigl(\pi\,\tfrac{s - s_{\mathrm{w}}}{S - s_{\mathrm{w}}}\bigr)\bigr)
      & \text{if } s \geq s_{\mathrm{w}} \;\text{and schedule}=\textsc{cosine-anneal},
  \end{cases}
  \label{eq:lambda_schedule}
\end{equation}
with base weight $\lambda_{0}$, warm-up length $s_{\mathrm{w}}$,
and total training horizon $S$. The motivation is empirical and
theoretical at once. Empirically: at initialisation the audio
representation has not yet been shaped by the flow-matching loss,
so its intra-batch geometry is essentially noise, and forcing it
to match the video geometry from step zero pulls the network into
degenerate solutions that satisfy the GW term at the cost of the
generative objective. Theoretically: existing fused-OT work (e.g.\
\citet{chen_2020_graph}) fixes the mixing $\alpha$ but not its
relative weight against the task loss. The schedule
\eqref{eq:lambda_schedule} resolves the latter dimension explicitly
and is, to our knowledge, novel for OT-based cross-modal
regularization.

\subsection{Why this isolates the relational hypothesis}
\label{sec:methodology:isolation}

Three properties make the construction above a clean test of the
thesis question. First, $\LFM$ in \eqref{eq:fm_loss} and the
backbone of \citet{cheng_2025_mmaudio} together constitute a
\emph{pointwise} alignment baseline: the generative target is
defined token by token via the constant velocity $x_1 - x_0$, and
all cross-modal information enters via mechanisms (joint attention,
adaLN with $c_g$, frame-aligned $c_f$) that operate on individual
tokens rather than on inter-token relations. The GW term
\eqref{eq:gw_value} is the smallest principled addition that
introduces an explicit \emph{relational} signal, because
Gromov--Wasserstein is, by construction, the OT object that
compares pairwise distance matrices rather than features.
Improvements traceable to $\LGW$ can therefore be attributed to
the relational signal specifically, rather than to extra capacity
or extra supervision in some other form.

Second, the four representation variants of
\S\ref{sec:methodology:reps} disentangle \emph{where} in the
network the relational signal is injected. \texttt{global}
operates before any learnable layer and asks whether the encoder
geometry already supports relational alignment without parameter
updates; \texttt{projected} injects the signal into the input
projections and lets it propagate through every transformer block;
\texttt{c\_g} aligns the very vector that drives adaLN modulation;
\texttt{fused} adds a Wasserstein term to test whether
relational and pointwise signals are complementary at the same
representation level; \texttt{got\_token} moves the GW objective
from cross-batch geometry to within-sample token geometry,
matching most closely the setting in which GW is typically deployed
\citep{chen_2020_graph,vayer_2019_optimal}. The five variants are
the five most natural injection points in MMAudio's
information flow; the ablation is exhaustive at this granularity.

Third, the schedule \eqref{eq:lambda_schedule} resolves a confound
that prior fused-OT work leaves implicit: improvements (or
regressions) cannot be attributed to a fortuitous choice of
$\lambda$ if the same schedule is used across all variants, and
ablating the schedule itself separates relational signal from
training dynamics.

The contribution of this chapter, relative to the upstream
MMAudio repository, is therefore localised in three additions: a
self-contained GW solver
(\texttt{src/mmaudio/model/gw\_regularization.py}), the integration
hook in the training loop that selects the active subset, applies
the schedule, and combines the two losses, and the configuration
surface (\texttt{config/base\_config.yaml::gw\_regularization}) that
exposes the variant, the schedule, the entropic regularization
$\varepsilon$, the FGW mixing $\alpha$, the detach flag, and the
anti-collapse parameters as the axes of the empirical study in
\S\ref{sec:results}. The flow-matching backbone, the encoders, the
synchronization module, the data pipeline, and every other
architectural choice are inherited unchanged from
\citet{cheng_2025_mmaudio}; this isolates the regularizer as the
single independent variable.
