\section{Results}
\label{sec:results}

\emph{Placeholder — this chapter is filled in after the runs in
Sec.\,\ref{sec:experiments} complete. No results are yet
committed to the repository (\texttt{output/} and \texttt{logs/}
are empty); the \texttt{results/} directory will hold the final
CSVs.}

\subsection*{Planned structure}
The chapter will mirror the hypothesis list from
Sec.\,\ref{sec:experiments:hypotheses}. Each subsection pairs an
aggregate metric table with one figure that shows the result
beyond aggregation.
\begin{enumerate}[leftmargin=*]
  \item \textbf{Baseline vs best GW} (H1). Full 4-metric row.
    Populated from \texttt{python experiments/analysis.py table
    --runs output/gw\_baseline output/gw\_var\_*}.
  \item \textbf{D1 variant sweep} (H2). Table of
    $\{$\texttt{global, projected, c\_g, fused}$\}$, plus
    loss-vs-iteration curves
    (\texttt{experiments/analysis.py curves --runs
    output/gw\_var\_*}) showing both $\LFM$ and $\LGW$ over
    training.
  \item \textbf{D2 $\lambda$ sweep} (H3). Metric curves across
    $\lambda_{\mathrm{GW}} \in \{0.001,\ldots,0.1\}$ at
    \texttt{BEST\_VARIANT}. If H3 holds, the FD$_{\text{PaSST}}$
    curve is U-shaped.
  \item \textbf{D3 detach ablation} (H4). 2-row table
    \texttt{detach\_video} vs not.
  \item \textbf{D4 schedule ablation} (H5). 3-row table.
  \item \textbf{D5 GW\,$\times$\,Synchformer} (interaction).
    $2 \times 2$ table of DeSync and IB-score.
  \item \textbf{OOD generalization.} Baseline vs GW on the
    held-out 30-class split; does GW help on unseen classes?
  \item \textbf{Geometry diagnostics.} Effective rank and top-$k$
    singular spectrum of video- and audio-side projected
    representations, baseline vs GW
    (\texttt{experiments/analysis.py geometry}).
  \item \textbf{Coupling matrices.} Within-class / cross-class
    mean coupling heatmap for the top-20 VGGSound classes
    (\texttt{experiments/analyze\_couplings.py}), with a ranked
    CSV of class self-alignment.
\end{enumerate}

\subsection*{What will count as negative evidence}
The experiments are designed so that a failure can be read
directly off the tables rather than explained away. Specifically:
\begin{itemize}[leftmargin=*,topsep=2pt]
  \item H1 fails if no variant beats baseline on at least two
    metrics within the tolerance set in
    Sec.\,\ref{sec:experiments:hypotheses}.
  \item H2 fails if variants are indistinguishable given the
    spread of the baseline-to-best gap, i.e. representation
    choice does not matter.
  \item H3 fails if the $\lambda$-sweep is monotone in either
    direction: that would mean GW is either a uniform gain (in
    which case the schedule/detach story is moot) or a uniform
    loss.
  \item H4 and H5 fail silently if they are not statistically
    separable; we will report the winning cell and the absolute
    gaps rather than bolding without a threshold.
  \item If coupling heatmaps are uniform (no block-diagonal
    structure) regardless of aggregate metrics, GW is solving its
    own objective without producing interpretable class-level
    alignment --- an important finding independent of the
    aggregate metric story.
\end{itemize}
